11th International Satisfiability Modulo Theories Competition (SMT-COMP 2016): Rules and Procedures

### 2016-04-21 rules/rules16.tex
Matthias Heizmann added as co-organizer. Minor edits.

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\texttt{Sylvain.Conchon@lri.fr}}} \\
\and
David D{\'e}harbe \\
Clearsy, France \& Federal University\\ of Rio Grande do Norte, Brazil \\
{\small\href{mailto:david.deharbe@clearsy.com}{\texttt{david.deharbe@clearsy.com}}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\texttt{tjark.weber@it.uu.se}}} \\
\\
}

@@

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Albert Oliveras, Aaron Stump, and Tjark Weber.} describes the rules
and competition procedures for SMT-COMP~2016.  The principal changes
from the previous competition rules are the following:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used in the 2015 SMT
  Competition.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, the SMT
  Competition has had sufficient computational resources to evaluate
  all solver entrants on all eligible benchmarks.  As a side-effect,
  the weighting of benchmark families that was achieved in early
  SMT-COMPs through a careful selection of benchmarks (based on
  benchmark difficulty and category) was lost.  We believe that a
  careful weighting (see Section~\ref{sec:division-scoring}), which
  de-emphasizes large benchmark families, will lead to more meaningful
  competition results.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  previous SMT-COMPs (except in some divisions that had a very small
  number of industrial benchmarks).  Therefore, industrial performance
  has very limited significance, and we consider the overhead to
  determine and report it separately not justified.
\end{itemize}

@@

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.5, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB commands that
satisfies the following requirements:
%
\begin{enumerate}
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command.  Note that \texttt{success} outputs are
  ignored by the post-processor used by the
  competition.\footnote{SMT-LIB~2.5 requires solvers to produce a
    \texttt{success} answer after each \akey{set-logic},
    \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
    (among others), unless the option \akey{:print-success} is set to
    false.  Ignoring the \texttt{success} outputs allows for
    submitting fully SMT-LIB~2.5 compliant solvers without the need
    for a wrapper script, while still allowing entrants of previous
    competitions to run without changes.}
\item In the application track, the \akey{:print-success} option must
  not be disabled.  The trace executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands.
\item The script may contain any number of $\akey{set-info}$ commands.
  These may be ignored by solvers.
\item The script may contain any number of $\akey{declare-sort}$
  commands.  All sorts declared with a $\akey{declare-sort}$ command
  must have zero arity.
\item Application track scripts may contain $\akey{define-sort}$
  commands.  \textcolor{red}{TODO: Some main track benchmarks also
    appear to contain \akey{define-sort} and \akey{define-fun}
    commands.}
\item The script may contain any number of $\akey{declare-fun}$
  commands.  All formulas in the script belong to the benchmark's
  logic, with any free symbols declared in the script.
\item In the main track, there is exactly one $\akey{check-sat}$
  command, following possibly several $\akey{assert}$ commands.
\item In the application track, there are one or more \akey{check-sat}
  commands, each preceded by one or more \akey{assert} commands.
  There may also be zero or more \akey{push 1} commands, and zero or
  more \akey{pop 1} commands, consistent with the use of those
  commands in the SMT-LIB standard.
\item The script may optionally contain an $\akey{exit}$ command as
  its last command.  In the application track, this command must not
  be omitted.
\item No other commands besides the ones just mentioned may be used.
\item Named formulas are not used in benchmark scripts.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Standard''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

For a given competition benchmark~$b$, let~$F_b \geq 1$ be the total
number of benchmarks in $b$'s benchmark family that were used in the
competition.  We define the weight for benchmark~$b$ as $\alpha_b = (1
+ \log F_b) / F_b$.\footnote{See Section~\ref{sec:competition-wide}
  for a motivating discussion of log scaling.}  We define the
\emph{normalized weight} for benchmark~$b$ as $\alpha'_b = \alpha_b /
(\sum_{b'} \alpha_{b'})$, where the sum is over all benchmarks in the
division.

@@

We define a competition-wide metric for the main track, separately for
parallel and sequential performance, as follows.  Let $N_i$ be the
total number of benchmarks in division~$i$ that were used in the
competition, and let $\langle e_i, n_i, w_i, c_i\rangle$ be a solver's
raw score for this division (Section~\ref{sec:division-scoring}).  The
solver's competition-wide raw score is $\sum_i (e_i == 0 \,?\,
(n_i/N_i)^2 : -e_i) \log N_i$, where the sum is over all competitive
divisions into which the solver was
entered.\footnote{\emph{Rationale:} This metric purposely emphasizes
  breadth of solver participation---a solver participating in many
  logics need not be the best in any one of them.  The use of the
  square in the metric limits that somewhat---a solver still needs to
  do reasonably well compared to winners to be able to catch up by
  breadth of participation.  The non-linear metric also gives added
  weight to completing close to all benchmarks in a division.  The log
  scaling is a (somewhat arbitrary) means to adjust the scores for the
  wide variety of numbers of benchmarks in different divisions.  It
  seems a reasonable compromise between linearly combining numbers of
  benchmarks, which would overweigh large divisions, and simply
  summing the fraction of benchmarks solved, which would overweigh
  small divisions.  The metric is also quite simple, and the metric
  for a solver is independent of the performance of other solvers.
  Time is omitted from the metric because it is only of third
  importance in the regular competition metric, and is difficult to
  compare across divisions.}  The solver's competition-wide sequential
score is computed from its sequential division scores
(Section~\ref{sec:division-scoring}) according to the same formula.
We will recognize the best three solvers according to these metrics.

@@

SMT-COMP 2016 is organized under the direction of the SMT Steering
Committee. The organizing team is
%
\begin{itemize}
\setlength{\itemsep}{0pt}
\item \href{https://www.lri.fr/~conchon/}{Sylvain Conchon}~--
  Paris-Sud University, France (co-organizer)
\item \href{https://www.sites.google.com/site/deharbe/}{David
  D{\'e}harbe}~-- Clearsy, France, and Federal University of Rio
  Grande do Norte, Brazil (co-organizer)
\item \href{http://user.it.uu.se/~tjawe125/}{Tjark Weber}~-- Uppsala
  University, Sweden (chair)
\end{itemize}
%
Tjark Weber is responsible for policy and procedure decisions, such as
these rules, with input from the co-organizers. He is not associated
with any group creating or submitting solvers.

Many others have contributed benchmarks, effort, and feedback.  Clark
Barrett
% processed the newly submitted benchmarks.
is maintaining the SMT-LIB benchmark library.  The competition uses
the \href{https://www.starexec.org/}{StarExec} service, which is
hosted at the \href{http://www.cs.uiowa.edu/}{University of Iowa}.
Aaron Stump is providing essential StarExec support.

David D{\'e}harbe is associated with the solver group producing the
veriT solver.

@@

%% Local Variables:
%% mode: latex
%% mode: flyspell
%% ispell-local-dictionary: "american"
%% LocalWords: Conchon logics satisfiability SMT Sylvain Tjark
%% End:

>>>>

\author{%
Sylvain Conchon \\
Paris-Sud University \\
France \\
{\small\href{mailto:Sylvain.Conchon@lri.fr}{\texttt{Sylvain.Conchon@lri.fr}}} \\
\and
David D{\'e}harbe \\
Clearsy, France \& Federal University\\ of Rio Grande do Norte, Brazil \\
{\small\href{mailto:david.deharbe@clearsy.com}{\texttt{david.deharbe@clearsy.com}}} \\
\and
Matthias Heizmann \\
University of Freiburg \\
Germany \\
{\small\href{mailto:heizmann@informatik.uni-freiburg.de}{\texttt{heizmann@informatik.uni-freiburg.de}}} \\
{\small\textsc{(Unsat Core Track)}} \\
\and
Tjark Weber \\
Uppsala University \\
Sweden \\
{\small\href{mailto:tjark.weber@it.uu.se}{\texttt{tjark.weber@it.uu.se}}} \\
\\
}

@@

The rest of this document, revised from the previous
version,\footnote{Earlier versions of this document include
  contributions from Clark Barrett, Roberto Bruttomesso, David Cok,
  Sylvain Conchon, David D{\'e}harbe, Morgan Deters, Alberto Griggio,
  Albert Oliveras, Aaron Stump, and Tjark Weber.} describes the rules
and competition procedures for SMT-COMP~2016.  The principal changes
from the previous competition rules are the following:
\begin{itemize}
\item Benchmarks will use (a subset of) version~2.5 of the SMT-LIB
  language.  \emph{Rationale:} This is the latest version of the
  SMT-LIB language.  It was released on 2015-06-28, and is largely
  backwards-compatible to version~2.0, which was used for SMT-COMP
  2015.
\item Solver output may affect the score even when the solver does not
  terminate within the time limit.  (In 2015, main track solver output
  was ignored if the solver subsequently timed out.)  Solvers should
  take care not to accidentally produce output that contains
  \texttt{sat} or \texttt{unsat} even when they are killed.
  \emph{Rationale:} Users are likely to trust solver responses even
  when the solver continues to run for some time.
\item Divisions for floating-point arithmetic are no longer
  experimental, and will be considered competitive if the necessary
  requirements (see Section~\ref{sec:scoring}) are met.
  \emph{Rationale:} Floating-point divisions were experimental in
  2015.  By now, their definition is sufficiently stable, and they are
  supported by several solvers.
\item Division scores will be based on a weighted sum of scores for
  benchmark families.  \emph{Rationale:} For some years now, SMT-COMPs
  have had sufficient computational resources to evaluate all solver
  entrants on all eligible benchmarks.  As a side-effect, the
  weighting of benchmark families that was achieved in early SMT-COMPs
  through a careful selection of benchmarks (based on benchmark
  difficulty and category) was lost.  Since there are vast differences
  in size between benchmark families, we believe that a weighting that
  de-emphasizes large benchmark families (see
  Section~\ref{sec:division-scoring}) will lead to more meaningful
  competition results.
\item Best industrial performance will no longer be recognized
  separately.  \emph{Rationale:} While there is agreement in the SMT
  community to emphasize problems that come from real applications,
  industrial performance largely coincided with overall performance in
  SMT-COMP 2015.  We consider the effort to determine and report it
  separately no longer justified.
\end{itemize}

@@

The benchmark script is in the concrete syntax of the SMT-LIB format,
version~2.5, though with a restricted set of commands.  A benchmark
script is a text file containing a sequence of SMT-LIB commands that
satisfies the following requirements:
%
\begin{enumerate}
\item In the main track, there may be a (single) \akey{set-option
  :print-success ...} command.  Note that \texttt{success} outputs are
  ignored by the post-processor used by the
  competition.\footnote{SMT-LIB~2.5 requires solvers to produce a
    \texttt{success} answer after each \akey{set-logic},
    \akey{declare-sort}, \akey{declare-fun} and \akey{assert} command
    (among others), unless the option \akey{:print-success} is set to
    false.  Ignoring the \texttt{success} outputs allows for
    submitting fully SMT-LIB~2.5 compliant solvers without the need
    for a wrapper script, while still allowing entrants of previous
    competitions to run without changes.}
\item In the application track, the \akey{:print-success} option must
  not be disabled.  The trace executor will send an initial
  $\akey{set-option :print-success true}$ command to the solver.
\item The (single) $\akey{set-logic}$ command setting the benchmark's
  logic is the first command after any $\akey{set-option}$ commands.
\item The script may contain any number of $\akey{set-info}$ commands.
  These may be ignored by solvers.
\item The script may contain any number of $\akey{declare-sort}$
  commands.  All sorts declared with a $\akey{declare-sort}$ command
  must have zero arity.
\item The script may contain any number of $\akey{define-sort}$
  commands.  All sorts defined with a $\akey{define-sort}$ command
  must have zero arity.
\item The script may contain any number of $\akey{declare-fun}$
  commands.
\item The script may contain any number of $\akey{define-fun}$
  commands.
\item The script may contain any number of $\akey{assert}$ commands.
  All formulas in the script belong to the benchmark's logic, with any
  free symbols declared in the script.
\item Named formulas are not used in benchmark scripts.
\item In the main track, there is exactly one $\akey{check-sat}$
  command.
\item In the application track, there are one or more \akey{check-sat}
  commands.  There may also be zero or more \akey{push 1} commands,
  and zero or more \akey{pop 1} commands, consistent with the use of
  those commands in the SMT-LIB standard.
\item The script may optionally contain an $\akey{exit}$ command as
  its last command.  In the application track, this command must not
  be omitted.
\item No other commands besides the ones just mentioned may be used.
\end{enumerate}
%
The SMT-LIB format specification is available from the ``Standard''
section of the SMT-LIB website~\cite{SMT-LIB}.  Solvers will be given
formulas only from the divisions into which they have been entered.

@@

For a given competition benchmark~$b$, let~$F_b \geq 1$ be the total
number of benchmarks in $b$'s benchmark family that were used in the
competition.  We define the weight for benchmark~$b$ as $\alpha_b = (1
+ \log_e F_b) / F_b$.\footnote{See Section~\ref{sec:competition-wide}
  for a motivating discussion of log scaling.}  We define the
\emph{normalized weight} for benchmark~$b$ as $\alpha'_b = \alpha_b /
(\sum_{b'} \alpha_{b'})$, where the sum is over all benchmarks in the
division.

@@

We define a competition-wide metric for the main track, separately for
parallel and sequential performance, as follows.  Let $N_i$ be the
total number of benchmarks in division~$i$ that were used in the
competition, and let $\langle e_i, n_i, w_i, c_i\rangle$ be a solver's
raw score for this division (Section~\ref{sec:division-scoring}).  The
solver's competition-wide raw score is $\sum_i (e_i == 0 \,?\,
(n_i/N_i)^2 : -e_i) \log_e N_i$, where the sum is over all competitive
divisions into which the solver was
entered.\footnote{\emph{Rationale:} This metric purposely emphasizes
  breadth of solver participation---a solver participating in many
  logics need not be the best in any one of them.  The use of the
  square in the metric limits that somewhat---a solver still needs to
  do reasonably well compared to winners to be able to catch up by
  breadth of participation.  The non-linear metric also gives added
  weight to completing close to all benchmarks in a division.  The log
  scaling is a (somewhat arbitrary) means to adjust the scores for the
  wide variety of numbers of benchmarks in different divisions.  It
  seems a reasonable compromise between linearly combining numbers of
  benchmarks, which would overweigh large divisions, and simply
  summing the fraction of benchmarks solved, which would overweigh
  small divisions.  The metric is also quite simple, and the metric
  for a solver is independent of the performance of other solvers.
  Time is omitted from the metric because it is only of third
  importance in the regular competition metric, and is difficult to
  compare across divisions.}  The solver's competition-wide sequential
score is computed from its sequential division scores
(Section~\ref{sec:division-scoring}) according to the same formula.
We will recognize the best three solvers according to these metrics.

@@

SMT-COMP 2016 is organized under the direction of the SMT Steering
Committee. The organizing team is
%
\begin{itemize}
\setlength{\itemsep}{0pt}
\item \href{https://www.lri.fr/~conchon/}{Sylvain Conchon}~--
  Paris-Sud University, France (co-organizer)
\item \href{https://www.sites.google.com/site/deharbe/}{David
  D{\'e}harbe}~-- Clearsy, France, and Federal University of Rio
  Grande do Norte, Brazil (co-organizer)
\item
  \href{https://swt.informatik.uni-freiburg.de/staff/heizmann}{Matthias
    Heizmann}~-- University of Freiburg, Germany (co-organizer, unsat
  core track)
\item \href{http://user.it.uu.se/~tjawe125/}{Tjark Weber}~-- Uppsala
  University, Sweden (chair)
\end{itemize}
%
Tjark Weber is responsible for policy and procedure decisions, such as
these rules, with input from the co-organizers. He is not associated
with any group creating or submitting solvers.

Many others have contributed benchmarks, effort, and feedback.  Clark
Barrett is maintaining the SMT-LIB benchmark library.  The competition
uses the \href{https://www.starexec.org/}{StarExec} service, which is
hosted at the \href{http://www.cs.uiowa.edu/}{University of Iowa}.
Aaron Stump is providing essential StarExec support.

David D{\'e}harbe is associated with the solver group producing the
veriT solver.  Matthias Heizmann is associated with the solver group
producing the SMTInterpol solver.

@@

%% Local Variables:
%% mode: latex
%% mode: flyspell
%% ispell-local-dictionary: "american"
%% LocalWords: arity Conchon logics satisfiability SMT Sylvain Tjark
%% End:
